A distributed key-value store using Raft

### 2014-06-04 proj/paper.tex
more paper

However, if a fault occurs during this process which leaves the leader unable to contact a majority of nodes, no response may be generated. To handle this case, the client nodes associate a $300$ms timeout with each request, and will retry the request if this timeout expires. The value of the timeout was chosen to be small while still being long enough to wait out leadership changes. After some number of retries (I picked $5$ to allow for two cycles of timeout-redirects), the client will return an error to the broker. 
\subsection{SET requests}
The beginning and end of a \verb|SET| request look much like those of \verb|GET| requests, wherein the broker submits a request to the client, the client repackages it for the cluster, and the response follows the same chain backwards. The middle part of a \verb|SET| request differs, as setting a value requires a change to be committed into the replicated state. To do this, the leader appends an entry into its log containing the key, the value, the request id, and the client's name. It then runs through the Raft consensus algorithm until the entry is committed. In this implementation, committing a \verb|SET| request entry on any node involves writing it to the replicated state, then messaging the client recorded in the entry to let it know that the request has been completed. This results in all nodes sending the same success message to the client, which improves the odds that the client will see one of them in the event of a network partition. However, it also means that the client has to deal with filtering out duplicates. It does this by using the unique request id, which is included in all messages sent from and to the client.
\end{document}

>>>>

However, if a fault occurs during this process which leaves the leader unable to contact a majority of nodes, no response may be generated. To handle this case, the client nodes associate a $600$ms timeout with each request, and will retry the request if this timeout expires. The value of the timeout was chosen to be small while still being long enough to wait out leadership changes. After some number of retries (I picked $5$ to allow for two cycles of timeout-redirects), the client will return an error to the broker. 
\subsection{SET requests}
The beginning and end of a \verb|SET| request look much like those of \verb|GET| requests, wherein the broker submits a request to the client, the client repackages it for the cluster, and the response follows the same chain backwards. The middle part of a \verb|SET| request differs, as setting a value requires a change to be committed into the replicated state. To do this, the leader appends an entry into its log containing the key, the value, the request id, and the client's name. It then runs through the Raft consensus algorithm until the entry is committed. In this implementation, committing a \verb|SET| request entry on any node involves writing it to the replicated state, then messaging the client recorded in the entry to let it know that the request has been completed. This results in all nodes sending the same success message to the client, which improves the odds that the client will see one of them in the event of a network partition. However, it also means that the client has to deal with filtering out duplicates. It does this by using the unique request id, which is included in all messages sent from and to the client.

The cluster may fail to respond to \verb|SET| requests as with \verb|GET| requests. However, in the case of \verb|SET|, retrying may lead to duplicate entries being committed into the replicated state. This is an acceptable outcome -- the last \verb|SET| entry will effectively override any earlier ones, and since the last \verb|SET| entry must be committed before the client reports success back to the broker, we can simply choose the commit of the last \verb|SET| entry to be the linearization point of the request from the broker.
\subsection{Fault tolerance}
Due to the majority-based voting of the Raft internals, this implementation is able to handle the failure of a minority of its nodes while being able to return to a state where it available. With respect to consistency, the original paper by Ongaro and Ousterhout gives a list of consistency properties in Figure 3 which have been shown to hold at all times. However, this only implies that this implementation is only eventually consistent. Consider a run where the leader has committed some key (and therefore sent notification to the client). If the leader fails between now and when it next contacts its followers, none of its followers will have committed the entry, so when the next leader comes to power, it will not have committed the entry either. If the client issues a \verb|GET| request at this point, the leader may not have determined that the entry was safe to commit before responding to the \verb|GET|, so the client may receive stale data. However, as the algorithm runs, the entry will eventually be committed due to the Leader Completeness property, guaranteeing eventual consistency. Unfortunately, since election timeouts are randomized and the broker can't perform commands conditioned on the contents of messages, I can't come up with a script that demonstrates this happening.

With respect to availability, it's easy to see how Raft may become unavailable under particular circumstances. If luckily timed transient node failures keep hitting the leaders just after they are elected, Raft will be unavailable. In addition, dueling candidates during elections can also impact availability. However, this is a rare case, so in most circumstances with isolated single failures, availability will only be impacted for a fraction of a second as leader election takes place. Thus, if we view this implementation in the CAP framework, it sits somewhere in the middle. It guarantees eventual consistency (but probably fully consistent without the lucky timing described above), and if we view error messages as the equivalent of not completing in a finite time, it has availability in the absence of really luckily timed faults.
\section{Script descriptions}
First, before I start describing the scripts, there is a small caveat. Since the elections are randomized to some extent, there can be a pretty long setup period in the beginning where there are no leaders. If this happens and if the first thing in the broker script is a set, then everything later on goes awry as well. Thus, everything I'm describing here assumes speedy elections, which happen most of the time but not always.
\subsection{Leader failure}
This script is named \verb|leader_failure.chi|, and it illustrates the scenario where the leader experiences a temporary network partition that isolates it from the rest of the network. The desired action in this case is for the four remaining followers to elect a new leader and keep serving requests, and upon rejoining, the former leader will realize that it is out-of-date, revert to a follower, and catch up on its log.

The logging output of a sample run is attached as \verb|leader_failure.out|. At line 76, the client first receives a \verb|SET| request from the broker and forwards it to \verb|A1|, who happens to be the leader. It then broadcasts AppendEntries messages, and \verb|A2| and \verb|A3| seem to have gotten them first (lines 80, 84). Having received affirmation from a majority, \verb|A1| goes ahead and commits (line 129, the ordering is kinda messed up), and the broker receives a response (127). Next, the broker issues a \verb|GET|, which \verb|A1| responds to immediately after the next heartbeat (129, 136). The broker then creates a partition, and $151$ms later, \verb|A3| times out. In the subsequent lines, \verb|A2| and \verb|A4| grant votes to \verb|A3|, while \verb|A5| fails to grant a vote due to an incompatibility in its log (probably from starting so late). In any case, \verb|A3| receives a majority and becomes leader (165). Meanwhile, the client has timed out contacting \verb|A1| for the \verb|GET|, as \verb|A1| is no longer able to receive heartbeat responses (195). Thus, it picks a random node to try (\verb|A2|) and gets redirected to \verb|A3| (200). \verb|A3| is then able to respond to the \verb|GET| request as leader, and similarly with the subsequent \verb|SET|.

Later, at line 235, the network partition is resolved, and \verb|A1| is able to contact other nodes again. Immediately, it sends out a heartbeat and is rejected by \verb|A5| because its term is lagging (237). \verb|A1| then receives AppendEntries from the current leader \verb|A3|, rejects them a few times because it is too out-of-date, and finally succeeds in appending 3 entries (241, 249, 253). Meanwhile, \verb|A3| handles the \verb|GET| request from the client in routine fashion, without consulting with the out-of-date \verb|A1|.
\end{document}
